---
title: "p8105_hw1_hc3448"
output: github_document
date: "2024-09-16"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Problem 1

```{r}
data("penguins", package = "palmerpenguins")
```


The `penguins` dataset contains `r nrow(penguins)` rows and `r ncol(penguins)` columns. It provides information on penguin species (value: `r paste(unique(penguins$species), collapse = ', ')`), on islands (value: `r paste(unique(penguins$island), collapse = ', ')`), and with other important variables such as `sex`, `year`, `bill_depth_mm`, `flipper_length_mm`, and `body_mass_g`. The mean flipper length of the penguins is `r mean(penguins$flipper_length_mm, na.rm = TRUE)` mm.

```{r message =FALSE, warning = FALSE}
library(ggplot2)
ggplot(data = penguins, aes(x = bill_length_mm, y = flipper_length_mm, color = species)) +
  geom_point() +
  labs(title = "Flipper Length vs Bill Length by Penguin Species",
       x = "Bill Length (mm)",
       y = "Flipper Length (mm)") +
  theme_minimal()

ggsave("penguin_scatterplot.png", width = 8, height = 6)
```

## Problem 2

```{r message = FALSE}
library(tidyverse)
```

```{r}
set.seed(1)

df <- tibble(
  numeric_var = rnorm(10),  #a random sample of size 10 from a standard Normal distribution             
  logical_var = rnorm(10) > 0, #a logical vector indicating whether elements of the sample are greater than 0   
  char_var = letters[1:10],   #a character vector of length 10
  factor_var = factor(rep(c("A", "B", "C"), length.out = 10))  #a factor vector of length 10, with 3 different factor “levels”
)

# Extract columns using pull() and calculate means
mean_numeric <- df %>% pull(numeric_var) %>% mean()
mean_logical <- df %>% pull(logical_var) %>% mean()
mean_char <- df %>% pull(char_var) %>%  mean()
mean_factor <- df %>% pull(factor_var) %>%  mean()

list(mean_numeric = mean_numeric,
     mean_logical = mean_logical,
     mean_char = mean_char,
     mean_factor = mean_factor)
```

The mean calculation of the random sample of size 10 from a standard Normal distribution works directly as it is a numeric vector.   
The mean calculation of the logical vector works because true is treated as 1 and false as 0. The mean will be the proportion of true values in the logical vector.  
The mean calculation of the character vector doesn't work.  
The mean of the factor vector doesn't work because it represents categorical data and not numercial values.  

```{r warning=FALSE, results = FALSE}
mean_logical <- df %>% pull(logical_var) %>% as.numeric() %>% mean()
mean_char <- df %>% pull(char_var) %>% as.numeric() %>% mean()
mean_factor <- df %>% pull(factor_var) %>% as.numeric() %>% mean()

list(mean_logical = mean_logical,
     mean_char = mean_char,
     mean_factor = mean_factor)

```

Logical: TRUE becomes 1 and FALSE becomes 0, so you get a numeric vector with these values.  
Character: Characters are not directly convertible to numeric. It will return NA with a warning unless the characters are valid numeric strings.  
Factor: The underlying integer codes of the factor levels are converted to numeric. If the factor levels are A, B, C, the conversion will yield the integers corresponding to these levels.

Yes. When we attempt to take the mean of a character or factor variable without converting them to numeric first, we get an error because the mean function expects numeric input. Converting factors and characters to numeric makes sense because it allows you to compute a mean based on their underlying numeric representation.
